---
title: "Identifying Clusters in a Spatial Point Pattern using Voronoi Tessellations"
author: "Tejal Malde"
date: "29/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(spatstat)
library(deldir)
```


## Code from Chapter 2 - Using Voronoi Diagrams to test for Clustering

Function that, given an area size and intensity, will run a Poisson process and output the normalised areas for the tiles of the Voronoi Tessellation:

```{r Poisson Process}
CSR_areas <- function(A, lambda){
  x.len <- sqrt(A)
  y.len <- sqrt(A)

  pointpr <- rpoispp(lambda, win = owin(c(0,x.len),c(0,y.len)), nsim = 1)
  areas <- dirichletAreas(pointpr)
  areas.norm <- areas/(mean(areas))
  return(areas.norm)  
}   
```

And similarly for a Thomas Process:

```{r Thomas Process}
Clustered_areas <- function(A, lambda_p, eta, sd){
  x.len <- sqrt(A)
  y.len <- sqrt(A)

  pointpr <- rThomas(lambda_p, sd, eta, win = owin(c(0,x.len),c(0,y.len)), nsim = 1)
  pointpr.vor <- dirichlet(pointpr)
  areas <- tile.areas(pointpr.vor)

  areas.norm <- areas/(mean(areas))
  return(areas.norm)
}   
```

We can then use this code to calculate the p values and run the hypothesis test, summing the number of times we reject the null.

```{r}
p.arr <- array(rep(0,10*10*100),c(10,10,100))
for (A in seq(5,50,by=5)) {
  for (lam in seq(5,50,by=5)) {
    for (i in 1:100) {
      a1 <- CSR_areas(A,lam)
      p.arr[(lam/5),(A/5),1] <- ks.test(a1,"pgamma",3.5,3.5)$p.value
    }
  }
}

p.graph <- array(rep(0,10*10),c(10,10))
for (k in 1:10) {
  for (j in 1:10) {
    p.graph[k,j] <- sum(p.arr[k,j,] < 0.05)
  }
}
```

To speed up the process of calculating the areas, I simply exported the point pattern and imported the areas, on which I ran my KS test.

To produce a QQ-Plot:

```{r}
A = 50
lam = 20
a1 <- CSR_areas(A,lam)
a2 <- (qgamma((0:99)/100,3.5,rate=3.5))
qqplot(a1,a2,pch=3,main=expression(paste('QQ-Plot: A=50, ',lambda,'=20')),xlab='Empirical Areas',ylab='Theoretical Distribution')
lines(0:3,0:3,col='darkgreen') # line x = y
```

## Code from Chapter 3 - Identifying Clusters with Voronoi Diagrams

Function that determines whether points are a cluster centre, border or not a cluster (labelled 2, 1, and 0 respectively):

```{r}
pointtype <- function(pointpr, pointpr.deldir, thresh){
  
  pointprtyp = rep(0, npoints(pointpr))

  for (i in 1:npoints(pointpr)){
    if (pgamma((pointpr.deldir$summary$dir.area[i])/(mean(pointpr.deldir$summary$dir.area)), 3.5, 3.5) < thresh){
      for  (j in 1:length(pointpr.deldir$dirsgs$ind1)){
        if (pointpr.deldir$dirsgs$ind1[j] == i){
          pointprtyp[pointpr.deldir$dirsgs$ind2[j]] = 1
       }
        if (pointpr.deldir$dirsgs$ind2[j] == i){
          pointprtyp[pointpr.deldir$dirsgs$ind1[j]] = 1
       }
      }
    }
  }
  for (i in 1:npoints(pointpr)){
    if (pgamma((pointpr.deldir$summary$dir.area[i])/(mean(pointpr.deldir$summary$dir.area)), 3.5, 3.5) < thresh){pointprtyp[i] = 2}}
  
  
return(pointprtyp)
}
```

Function that combines cluster centres (red tiles) together to give them the same cluster label.

```{r}
adjredpts <- function(index, pointpr.deldir, pointprtyp, parentgrp2){
  for (j in 1:npoints(pointpr)){
  if ((pointprtyp[j] == 2)&(parentgrp2[j] == 0)){
    if (any(pointpr.deldir$dirsgs$ind2[which(pointpr.deldir$dirsgs$ind1 == j)] == index)){
        parentgrp2[j] = parentgrp2[index]   
      }
    if (any(pointpr.deldir$dirsgs$ind1[which(pointpr.deldir$dirsgs$ind2 == j)] == index)){
        parentgrp2[j] = parentgrp2[index]
      }
   }
}
  return(parentgrp2)
}
```

Finally, putting this together, the function "label_clusters" takes in a point pattern with the coordinates of the corners of the rectangular window and outputs a list where the numbers correspond to which cluster each point is in. (0 means not in a cluster)

```{r}
label_clusters <- function(pointpr, x.min, x.max, y.min, y.max, thresh){
  pointpr.deldir <- deldir(pointpr, rw = (c(x.min,x.max, y.min,y.max)))
  pointprtyp <- pointtype(pointpr, pointpr.deldir, thresh)
  
  parentgrp <- numeric(npoints(pointpr))
  parentgrp1 <- parentgrp
  parentgrp2 <- parentgrp
  z = 1
  
  for (i in 1:npoints(pointpr)){
    if ((parentgrp2[i] == 0) & (pointprtyp[i] == 2)){
      parentgrp2[i] = z
      while (any(parentgrp2 != parentgrp)){
        parentgrp1 <- parentgrp2
        for (k in 1:length(parentgrp)){
          if (parentgrp[k] != parentgrp1[k]){
            parentgrp2 <- adjredpts(k, pointpr.deldir, pointprtyp, parentgrp2)
          }
        }
        parentgrp <- parentgrp1
      }
      z = z+1
    }
  }
  
  for (i in 1:npoints(pointpr)){
    if (pointprtyp[i] == 1){
      neighbours <- numeric()
      for (j in 1:npoints(pointpr)){
        if ((pointprtyp[j] == 2) & ((any(pointpr.deldir$dirsgs$ind2[which(pointpr.deldir$dirsgs$ind1 == j)] == i)) | (any(pointpr.deldir$dirsgs$ind1[which(pointpr.deldir$dirsgs$ind2 == j)] == i)) )){
          neighbours <- append(neighbours, parentgrp[j])
          
        }
      }
      parentgrp[i] <- neighbours[which.min(pointpr.deldir$summary$dir.area[neighbours])]

    }
  }
  for (i in (1:length(parentgrp))){
    if (is.na(parentgrp[i])){
      parentgrp[i] = 0
    }
  }
  return(parentgrp)
}
```

For example, when importing data (such as the challenge datasets) it could be done as follows:

```{r}
challenge_data <- read.csv('C:\\Users\\teji1\\OneDrive - Imperial College London\\UROP 2020\\Datasets Cluster Analysis\\Datasets_Cluster_Analysis_Showtime_2020\\01_challenge.csv', header=F)
x <- challenge_data[,1]
y <- challenge_data[,2]
pointpr <- ppp(x, y, window = owin(c(0,10000), c(0,10000)))
parentgroup <- label_clusters(chall_pp, 0, 10000, 0, 10000, 0.01)
```


